### 2018-08-13 main.tex
Update about posterior mean

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{mathrsfs}
\usepackage{float}
% ======    PACKAGES
\usepackage{slivkins-setup,slivkins-theorems}
%\usepackage{amsmath, amsfonts, amssymb, amsthm, amsbsy, amscd, bm, bbm}
\usepackage{amsbsy, amscd, bm, bbm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[small,bf]{caption}
\setlength{\captionmargin}{30pt}
\usepackage{subcaption}
\captionsetup[sub]{margin=10pt,font=small}
\usepackage{color}
\usepackage{ifthen}
\usepackage{xspace}
\usepackage[noend]{algorithmic}
\usepackage{algorithm}
\usepackage[colorlinks,citecolor={black},urlcolor={black},linkcolor={black}]{hyperref}
\usepackage{url}
\usepackage{tocbibind}
\usepackage{enumerate}
\usepackage{mdframed}
\usepackage{comment}

\DeclareMathOperator*{\argmin}{argmin}

% a very useful package for edits and comments, from David Kempe (USC)
\usepackage{color-edits}
%\usepackage[suppress]{color-edits}  % use this to suppress the package
\addauthor{as}{red}      % as for Alex
\addauthor{jm}{blue}     % jm for Jieming
\addauthor{ni}{green}     % ni for Nicole
\addauthor{sw}{magenta}     % sw for Steven
% e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{conjecture}[theorem]{Conjecture}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{remark}{Remark}[section]
%\newtheorem{claim}{Claim}[section]
%\newtheorem{example}{Example}[section]

\newcommand{\term}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\thread}{\term{thread}}

\newcommand{\DT}{\mD} % Alex' notation for distribution over types.

\def\DKL{\textbf{D}_{\term{KL}}}
\def\D{\mathbb{D}}
\def\E{\mathbb{E}}

\def\A{\mathcal{A}}
\def\M{\mathcal{M}}
\def\S{\mathcal{S}}
\def\X{\mathcal{X}}
\def\EX{\term{EX}}
\def\Ds{*}
\def\EE{\mathcal{E}}
\def\varTheta{\bold{\Theta}}
\def\varOmega{\bold{\Omega}}

\def\ALGG{Greedy }
\def\ALGPG{Parallel Greedy }
\def\2LEVEL{Two-level Policy}
\def\Ber{\emph{Ber} }

\def\pmn{\hat{\mu}}
\def\emn{\bar{\mu}}

\title{Bayesian Exploration with Unbiased History}

\author{
}
\begin{document}
\begin{titlepage}
\maketitle

\thispagestyle{empty}
\begin{abstract}
aaa
\end{abstract}
\end{titlepage}

\input{model}

\section{Social Learning}
\begin{lemma}
\ALGG of $T_c$ rounds pulls each arm w.p. at least $p_c$. 
\end{lemma}

\begin{corollary}
\ALGPG of $c \cdot T_c \cdot \frac{1}{p_c} \cdot \log(1/\delta)$ rounds pulls each at least once w.p. $1-\delta$. \jmcomment{Do we actually need this?}
\end{corollary}

\input{2level}

\input{3level}

\appendix

\input{posterior}
\end{document}