### 2018-08-30 prelim.tex
3-level proof

%!TEX root = main.tex
\section{Preliminaries}
\label{sec:prelim}

\begin{theorem}[Chernoff Bound]
Let $X_1,...,X_n$  be independent random variables such that $X_i \in [0,1]$ for all $i$. Let $\bar{X} = \frac{1}{n}\sum_{i=1}^n X_i$ denote their empirical mean. Then
\[
\Pr[ |\bar{X} - \E[\bar{X}]| > \varepsilon] \leq 2\exp(-2n\varepsilon^2).
\]
\end{theorem}

\begin{theorem}[Berry-Essen Theorem]
Let $X_1,...,X_n$ be i.i.d. variables with $\E[ (X_1 - \E[X_1])^2] = \sigma^2 >0$ and $\E[ |X_1 - \E[X_1]|^3] =\rho <\infty$. Let $\bar{X} = \frac{1}{n} \sum_{i=1}^n X_i$. Let $F_n$ be the cumulative distribution function of $\frac{(\bar{X} - \E[\bar{X}]) \sqrt{n}}{\sigma}$ and $\Phi$ be the cumulative distribution function of the standard normal distribution. For all $x$ and $n$,
\[
|F_n(x) - \Phi(x) | \leq \frac{\rho}{2\sigma^3\sqrt{n}}.
\]
\end{theorem}