### 2018-09-04 llevel.tex
Start >3 level proof

%!TEX root = main.tex

\section{$L$-level Recommendation Policy}
In this section, we design an $L$-level recommendation policy for $L > 3$. By having more than 3 levels, we get even smaller regret.

Our recommendation policy has $L$ levels and each level has $S^2$ groups for $S = \log(T)$. Label the groups in the $l$-th level as $G_{l,a,b}$ for $a,b \in [S]$.

We set the group size as following. For $l < L$,
\[
T_l = T^{\frac{2^{L-1} + 2^{L-2} + \cdots + 2^{L-l}}{2^{L-1}+ 2^{L-2} + \cdots + 1}}.
\]
and 
\[
T_L = T/S^2 - T_1 \cdot \GdT - T_2 - \cdots - T_{l-1}. 
\]
Each first-level group ($G_{1,a,b}$ for $a,b\in [S]$) has $T_1$ \ALGG runs in parallel. For $l \geq 2$, there are $T_l$ agents in group $G_{l,a,b}$. $T_L$ is a little bit different because we want total number of agents to be $T$.

Finally we define the information flow. Agents in the first level only observe the history defined in the \ALGG. For agents in group $G_{l,a,b}$ with $l\geq 2$, they observe all the history in the first $l-2$ levels and history in group $G_{l-1,b,c}$ for all $c \in [S]$. For agents which are not in the first level, if they are in the same group, they observe the same history.

\begin{theorem}
The $L$-level recommendation policy gets expected regret $O\left(T^{\frac{2^{L-1}}{2^L-1}} \log^2(T) L \right)$. In particular, if we pick $L = \log\log(T)$, then we get regret $O(T^{1/2} \log^2(T) \log\log(T))$.  
\end{theorem}

\begin{proof}

\end{proof}