### 2018-10-09 algg.tex
Hello

%!TEX root = main.tex
\section{\ALGG}
We first consider \ALGG which simply shows agents the full history. We discuss in this section the exploration of \ALGG under Assumption \ref{ass:embehave}.

\begin{lemma}
\label{lem:greedy}
There exist two constants $\GdT = O_K(1)$ and $\GdP = \Omega_K(1)$ such that for any arm $a$, with probability at least $\GdP$, \ALGG of $\GdT$ rounds pulls this arm at least once. \swcomment{maybe explain what $O_k$ means}
\end{lemma}

\begin{proof}
Set $\GdT = (K-1) \cdot c_T + 1$ and $\GdP = (1/3)^{\GdT}$. We fix the behavior of agents in $\GdT$ rounds and consider the case that all realized rewards in $\GdT$ rounds are 0.

In this case, we want to show that arm $a$ is pulled at least once. We prove by contradiction. Suppose arm $a$ is not pulled. By averaging argument, we know that there is some other arm $a'$ which is pulled at least $c_T + 1$ rounds. Suppose in round $t$, arm $a'$ is pulled exactly $c_T + 1$ times. By Assumption \ref{ass:embehave}, we know 
\[
\hat{\mu}_{a'}^t \leq 0 + c_m / \sqrt{c_T} < 1/3. 
\]
On the other hand, we have $\hat{\mu}_a^t \geq 1/3 > \hat{\mu}_{a'}^t$. This contradicts with the fact that in round $t$, arm $a'$ is pulled.

In addition, we know that this case happens with probability at least $(1/3)^{\GdT} = \GdP$ as each arm's mean is at most $2/3$.\swcomment{where did we assume that?} To sum up, we know that with probability at least $\GdP$, \ALGG of $\GdT$ rounds pulls arm $a$ at least once. 
\end{proof}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: