### 2018-10-16 algg.tex
rewriting

%!TEX root = main.tex
\section{\ALGG: Revealing Full History}

\swedit{Before we describe our algorithms, we will first analyze a
  simple policy \ALGG that fully discloses the history over all
  previous rounds, and lets the agents make the ``greedy'' choices
  across all rounds. Even though the algorithm does not guarantee low
  regret, it will be used as a subroutine in our main algorithms.  In
  particular, we show that because of the randomness in the rewards,
  with constant probability, $\ALGG$ will incentivize the agents to
  play each arm at least once.}

\begin{lemma}
\label{lem:greedy}
Under \Cref{ass:embehave}, there exist two constants $\GdT = O_K(1)$
and $\GdP = \Omega_K(1)$ such that for any arm $a$, with probability
at least $\GdP$, \ALGG of $\GdT$ rounds pulls this arm at least
once.\footnote{The notations $O_k$ and $\Omega_K$ hide the dependence
  on the number of arms $K$.}
\end{lemma}

\begin{proof}
  Set $\GdT = (K-1) \cdot c_T + 1$ and $\GdP = (1/3)^{\GdT}$. We fix
  the behavior of agents in $\GdT$ rounds and consider the case that
  all realized rewards in $\GdT$ rounds are 0.

In this case, we want to show that arm $a$ is pulled at least once. We prove by contradiction. Suppose arm $a$ is not pulled. By averaging argument, we know that there is some other arm $a'$ which is pulled at least $c_T + 1$ rounds. Suppose in round $t$, arm $a'$ is pulled exactly $c_T + 1$ times. By Assumption \ref{ass:embehave}, we know 
\[
\hat{\mu}_{a'}^t \leq 0 + c_m / \sqrt{c_T} < 1/3. 
\]
On the other hand, we have $\hat{\mu}_a^t \geq 1/3 > \hat{\mu}_{a'}^t$. This contradicts with the fact that in round $t$, arm $a'$ is pulled.

In addition, we know that this case happens with probability at least
$(1/3)^{\GdT} = \GdP$ as each arm's mean is at most $2/3$. To sum up,
we know that with probability at least $\GdP$, \ALGG of $\GdT$ rounds
pulls arm $a$ at least once.
\end{proof}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: