### 2018-11-02 llevel-overview.tex
editing...

%!TEX root = main.tex

\section{$\tilde O(\sqrt{T})$ regret with $L$-level policy}
\label{sec:llevel}
In this section, we give an overview of how we extend our three-level
policy to a more adaptive $L$-level policy for $L > 3$ in order to
achieve a regret rate of $O_K(\sqrt{T} \polylog(T))$. We provide two
such policies. The first policy achieves the root-$T$ regret rate with
$O(\log \log T)$ levels.

% In Theorem \ref{thm:llevel-1}, we show our $L$-level policy. We can
% achieve nearly optimal regret $O_K(T^{1/2} \polylog(T))$ with
% $O(\log\log(T))$ levels.
\begin{theorem}
\label{thm:llevel-1}
For any $L > 3$, there exists an $L$-level disclosure policy with
regret $$O_K\left(T^{2^{L-1}/(2^L-1)} \cdot \polylog(T) \right).$$ In
particular, there exists a $O(\log\log(T))$-level recommendation
policy with regret $O_K(T^{1/2} \polylog(T))$.
\end{theorem}

% In Theorem \ref{thm:llevel-2}, we show a policy with
Our second policy achieves an instance-dependent regret
guarantee. This policy has the same info-graph structure as the first
one in Theorem \ref{thm:llevel-1}, but requires a higher number of
levels $L = O(\log(T/\log\log(T)))$ and larger group sizes. We will
bound its regret as a function of the gap parameter $\Delta$ even
though the construction of the policy does not depend on $\Delta$. In
particular, this regret bound outperforms the one in Theorem
\ref{thm:llevel-1} when $\Delta$ is much bigger than $T^{-1/2}$.  It
also has the desirable property that the policy does not withhold too
much information from agents---any agent $t$ observes a good fraction
of history in previous rounds.

\begin{theorem}
\label{thm:llevel-2}
There exists a $O(\log(T)/\log\log(T))$-level policy such that for
every multi-armed bandit instance with gap parameter $\Delta$, the
policy has regret $$O_K(\min(1/\Delta, T^{1/2}) \cdot \polylog(T))$$
Moreover, under this policy, each agent $t$ observes a subhistory of
size at least $\Omega( t/\polylog(T))$.
\end{theorem}

Note for constant number of arms, this result mathches the optimal
regret rate (given in \Cref{eq:model-OptRegret}) for stochastic
bandits, up to logarithmic factors.

In this section, we present the main techniques in our solution, and
the full proofs of Theorem \ref{thm:llevel-1} and Theorem
\ref{thm:llevel-2} will be deferred to Appendix
\ref{sec:llevel-details}. Similarly as Section \ref{sec:3level}, we
first prove them in the case of 2 arms (Theorem \ref{thm:llevel} and
Corollary \ref{cor:llevel}). We then extend them to the case of
constant number of arms (Theorem \ref{thm:constarm}).

A natural idea to extend the three-level policy is to insert more
levels as multiple ``check points", so the policy can incentivize the
agents to perform more adaptive exploration. However, we need to
introduce two main modifications in the info-graph two accomendate
some new challenges. We will first informally describe our techniques
for the two-arm case.

% \begin{description}
% \item[Challenge 1] \emph{Exponentially growing number of groups.} 
% \item[Challenge 2] \emph{Uncertain number of arm pulls.}  
% \end{description}

\OMIT{We will now present the main idea of extending from 3-level to
  $L$-level is that instead of using the second level as one
  ``check-point'', we use more levels to have multiple ``check
  points''.  Some new challenges appear in this process. Here we give
  an overview of the additional techniques we use to get our $L$-level
  results.}

\paragraph{Interlacing connections between levels.}A tempted approach
to generalize the three-level policy is to build an $L$-level
info-graph with the structure of a $\NG$-ary tree: for every
$l\in \{2, \ldots , L\}$, each $l$-level group observes the
sub-history from a disjoint set $\NG$ groups in level $(l-1)$. The
disjoint sub-histories observed by all the groups in level $l$ are
independent, and under the small gap regime (similar to
Lemma~\ref{3levelsmallcase}) it ensures that each arm $a$ has a
``lucky'' $l$-level group of agents that only pull $a$. This ``lucky''
property is crucial for ensuring that both arms will be explored in
level $l$.

However, in this construction, the first level will have $\NG^{L-1}$
groups, which incurs $\NG^{\Omega(L)}$ regret. The exponential
dependence in $L$ will heavily limit the adaptivity of the policy, and
prevents having the number of levels for obtaining the result in
\Cref{thm:llevel-2}. To overcome this, we will design an info-graph
structure such that the number of groups at each level stays as
$\NG^2 = \Theta(\log^2(T))$.

We will leverage the following key observation: in order to maintain
the ``lucky'' property, it suffices to have the agents in different
$l$-th level groups to observe disjoint sub-histories that takes place
in level $(l - 1)$. Moreover, as long as the group size in levels
lower than $(l-1)$ are substantially smaller than group size of level
$l-1$, the ``lucky'' property does not break even if different groups
in level $l$ observe overlapping sub-history from levels
$\{1, \ldots, l-2\}$.

\OMIT{In our 3-level policy, the second level has
  $\NG = \Theta(\log(T))$ groups. We do so to ensure that in the small
  gap case (Lemma \ref{3levelsmallcase}), with high probability, both
  arms are pulled enough times in the second level. The argument
  relies on the fact that agents in different second-level groups
  observe disjoint histories of the first level and the independent
  randomness in first level groups guarantees each arm $a$ to have a
  ``lucky'' second-level group such that agents in that group all pull
  arm $a$ with high probability.

  Simply generalizing this idea to an $L$-level policy would give us a
  $\NG$-ary tree like structure and the first level will have
  $\NG^{L-1}$ groups. It incurs an extra $\NG^{\Omega(L)}$ factor in
  the regret. If we use $\log\log(T)$ levels as in Theorem
  \ref{thm:llevel-1}, this factor super poly-logarithmic. If we use
  $\log(T)/\log\log(T)$ levels as in Theorem \ref{thm:llevel-2}, this
  factor goes up to polynomial in $T$.

  In order to avoid this undesirable factor, we design a slightly
  different connecting structure between levels in our $L$-level
  policy.  The key observation is that, for any level
  $l \in \{3,\ldots,L-1\}$, we do want agents in different $l$-th level
  groups to see disjoint histories of the $(l-1)$-th level to ensure
  that there is some ``lucky'' group for each arm whose mean is close
  enough to the best arm. However, it does not matter much if pulls in
  levels below $l-1$ get observed by agents in different $l$-th level
  groups because group sizes in lower levels are much smaller than
  group sizes of level $l-1$ and the independent randomness in level
  $l-1$ is sufficient.}

This motivates the following interlacing connection structure between
levels. For each level in the info-graph, there are $\NG^2$ groups for
some $\NG = \Theta(\log(T))$. The groups in the $l$-th level are
labeled as $G_{l,u,v}$ for $u,v\in[\NG]$. For any $l \in \{2,\ldots,L\}$
and $u,v,w\in [\NG]$, agents in group $G_{l,u,v}$ see the history of
agents in group $G_{l-1,v,w}$ (and by transitivity all agents in
levels below $l-1$). See Figure \ref{fig:llevel-connecting} for a
visualization of simple case with $\NG = 2$). Two observations are
in order:
% \OMIT{(i) The number of groups does not grow with $L$ and stays as
%   $\NG^2 = \Theta(\log^2(T))$ in each level. Having these groups only
%   incurs a $\Theta(\log^2(T))$ multiplicative factor to the regret.\\}
\begin{itemize}
\item[(i)] Consider level $(l - 1)$ and fix the last group index to be
  $v$, and consider the set of groups
  $\cG_{l-1, v}=\{G_{l-1,i,v}\mid i \in [\NG]\}$ (e.g. $G_{l-1,1,1}$
  and $G_{l-1,2,1}$ circled in the Figure
  \ref{fig:llevel-connecting}). The agents in any group of
  $\cG_{l-1, v}$ observe the same sub-history. If the empirical mean
  of arm $a$ is sufficiently high in their shared sub-history, then all groups in $\cG_{l-1, v}$ will become ``lucky'' groups for $a$.

\item[(ii)] Every agent in level $l$ observes the sub-history from
  $\NG$ $(l-1)$-th level groups, each of which belonging to a
  different set $\cG_{l-1, v}$. Thus, for each arm $a$, we just need
  one set of groups $\cG_{l-1, v}$ in level $l-1$ to be ``lucky'' for
  $a$ and then all agents in level $l$ will see sufficient arm $a$
  pulls.
 \OMIT{For each set of groups
    in level $l-1$, agents in level $l$ see the history of agents in
    one of these groups. Therefore, for each arm $a$, we just need one
    set of groups in level $l-1$ to get ``lucky'' and then all agents
    in level $l$ will see enough pulls of arm $a$.}
\end{itemize}

\begin{figure}[H]
\centering
\begin{tikzpicture}  
 \foreach \x in {0,3,6,9}
 {
 \filldraw[fill=purple!20!white]
 (\x+0,7.5)--(\x+2,7.5)--(\x+2,8.5)--(\x+0,8.5)--cycle;
 \filldraw[fill=green!20!white]
 (\x+0,5)--(\x+2,5)--(\x+2,6)--(\x+0,6)--cycle;
 \filldraw[fill=blue!20!white]
 (\x+0,2.5)--(\x+2,2.5)--(\x+2,3.5)--(\x+0,3.5)--cycle;
 }
\foreach \y in {2.5,5,7.5}
{
  \node at (12.5,\y+0.5){$\cdots$};
}

\foreach \y in {3.5,6}
{
  \draw (1,\y) -- (1,\y+1.5);
  \draw (1,\y) -- (7,\y+1.5);

  \draw (4,\y) -- (1,\y+1.5);
  \draw (4,\y) -- (7,\y+1.5);

  \draw (7,\y) -- (4,\y+1.5);
  \draw (7,\y) -- (10,\y+1.5);

  \draw (10,\y) -- (4,\y+1.5);
  \draw (10,\y) -- (10,\y+1.5);
}

\foreach \u in {1,2}
{
	\foreach \v in {1,2}
	{
	\pgfmathsetmacro{\x}{((\u-1)*2+(\v-1))*3};
	\pgfmathsetmacro{\xa}{((\u-1))*3};
	\pgfmathsetmacro{\xb}{(2+(\u-1))*3};
   \node at(\x+1,8){$G_{l,\u,\v}$};
   \node at(\x+1,5.5){$G_{l-1,\u,\v}$};
   \node at(\x+1,3){$G_{l-2,\u,\v}$};
	}
}

  \node at(-1.2,3){\textbf{Level $l-2$}};
  \node at(-1.2,5.5){\textbf{Level $l-1$}};
  \node at(-1.2,8){\textbf{Level $l$}};
  \draw[->] (13.3,2)--(13.3,8.5);
  \node at(13.7,5)[ rotate=90]{Time};

 \draw [rounded corners=3mm, line width=1mm, red](-0.2,4.8)--(2.2,4.8)--(2.2,6.2)--(-0.2,6.2)--cycle;
  \draw [rounded corners=3mm, line width=1mm, red](5.8,4.8)--(8.2,4.8)--(8.2,6.2)--(5.8,6.2)--cycle;
\end{tikzpicture}
\caption{Connectiing structures between levels for the $L$-level policy.}
\label{fig:llevel-connecting}
\end{figure}

\paragraph{Amplifying groups for boundary cases.} Recall in our
3-level policy, the medium gap case (Lemma \ref{3levelmedium}) is the
boundary case when the gap $|\mu_1-\mu_2|$ is
$\Omega\left(\sqrt{\frac{1}{T_1}}\right)$ and
$O\left(\sqrt{\frac{\log(T)}{T_1}}\right)$. In this case, the gap is
not large enough to conclude that with high probability agents in both
the second level and the third level all pull the best arm. The gap is
also not small enough to conclude that with high probability both arms
are explored enough times in the second level. We need to worry about
the situation in which the sub-optimal arm does not get pulled enough
times in the first two levels and some agents in the third level may
pull the sub-optimal arm. Such situation is naturally ruled out in our
3-level policy for the following reason. Agents in the third level
observe the history of the entire first level while agents in the
second level only observes the history of a single first-level
group. For each arm, even if it is not explored enough times in the
second level, its empirical mean concentrates closer to its actual
mean in the history observed by a third-level agent than the one
observed by a second-level agent. Therefore, although the gap in the
medium gap case is just not large enough for agents in the second
level to all pull the best arm, it is large enough for agents in the
third level to all pull the best arm.

When we extend the 3-level policy to an $L$-level policy, we have such boundary case for each intermediate level. Moreover, the worry mentioned above does not get naturally resolved as in some of these boundary cases, the ratios between the upper and lower bounds of the gap rise from $\Theta(\sqrt{\log(T)})$ to $\Theta(\log(T))$. The reason for such rise is that, except the first level, we don't have good enough guarantee of the number of pulls of each arm. For example, in Figure \ref{fig:llevel-connecting}, when we talk about having enough arm $a$ pulls in the history observed by agents in $G_{l,1,1}$, it could be that  only agents in group $G_{l-1,1,1}$ are pulling arm $a$ and it also could be that most agents in groups $G_{l-1,1,1},G_{l-1,1,2},...,G_{l-1,1,\NG}$ are pulling arm $a$. Therefore our estimate of the number of arm $a$ pulls can be off by an $\NG=\Theta(\log(T))$ multiplicative factor. This finally makes the boundary cases harder to deal with.

In our $L$-level policy, after understanding the boundary cases well, we resolve this problem by introducing an additional type of groups: $\Gamma$-groups. For each $l \in [L], u,v \in [\NG]$, we create a $\Gamma$-group $\Gamma_{l,u,v}$. Agents in $\Gamma_{l,u,v}$ observe the same history as the one observed by agents in $G_{l,u,v}$ and the number of agents in $\Gamma_{l,u,v}$ is $\Theta(\log(T)$ times the number of agents in $G_{l,u,v}$. The main difference between the $G$-groups and $\Gamma$-groups is that the history of $\Gamma$ groups in level $l$ is not sent to agents in level $l+1$ but agents in higher levels. When we are in the boundary case in which we don't have good guarantees about the $l+1$ level agents' pulls, the new construction makes sure that agents in levels higher than $l+1$ get enough pulls of each arm and all pull the best arm. 
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: